---
title: "Mobile Device Usage & Behaviour Analysis"
author: Abigail Asamoah
output: html_notebook
date: 12-October-2024
---

### Import the required libraries below

```{r}
require(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
```

### Read file and show the column names and data types, and number of observations
```{r}
mob_data <- read.csv("../datasets/user_behavior_dataset.csv")

print("No of observations:")
str(mob_data)

summary(mob_data)
```
### Show the number of rows  and columns of data in the data set
```{r}

no_rows <- as.character(nrow(mob_data))
no_columns <- as.character(ncol(mob_data))

paste("There are", no_rows, " rows and", no_columns,"columns in the dataset.")

```
### Show the first 10 rows of the file for a quick review
```{r}
mob_data_rows <- head(mob_data, 10)
print(mob_data_rows)
```
### Review and find any missing values in the columns in the dataset
```{r}
colSums(is.na(mob_data))

```
### Review and find any missing values in the columns in the dataset

```{r}
duplicates <- mob_data[duplicated(mob_data),]
duplicates
```

### Que.1 Which operating System has the largest users? 
```{r}
#Grouping users by operating systems used
mob_data %>% 
  group_by(Operating.System) %>% 
  tally()

```


### Que.2 What are the types of devices we have across the two operating systems?
```{r}

library(dplyr)
mob_data %>% 
  group_by(Device.Model) %>% 
  tally()
```

### Que.3 What is the average screen time hours per day?
```{r}
print("The average screen hours in any given day:")
mean(mob_data$App.Usage.Time..min.day)

```

### Que.4 Calculate the proportion of iOS users and identify number of apps by device model
```{r}
mob_data %>%
  filter(Operating.System == "iOS") %>%
  group_by(Device.Model) %>%
  summarise(no_apps = (round(mean(Number.of.Apps.Installed), digits=0))) %>%
  arrange(desc(no_apps))

```

### Que.5 Calculate the proportion of Android users and identify number of apps by device model 
```{r}
mob_data %>%
  filter(Operating.System == "Android") %>%
  group_by(Device.Model) %>%
  summarise(no_apps = (round(mean(Number.of.Apps.Installed), digits=0))) %>%
  arrange(desc(no_apps))

```

### Que. 6 What is the total number of users by age group?
```{r}
mob_data["age_group"] = cut(mob_data$Age, c(0, 16, 35, 48, 64, Inf), c("0-16", "16-35", "35-48", "48-64", ">64"), include.lowest=TRUE)
mob_data %>% 
  group_by(age_group) %>% 
  tally()

```

### Que. 7 What is the total of hours spent on screen  by age group?
```{r}

mob_data %>% 
  group_by(age_group) %>%
  summarise(total_screen_time = (round(sum(Screen.On.Time..hours.day.), digits=0))) %>%
  arrange(desc(total_screen_time))


ggplot(mob_data, aes(x=age_group, y=Screen.On.Time..hours.day.)) +
 geom_bar(stat = "identity", width=0.4)

```


### Que. 8 What is the average battery drain time in a day for all device models?
```{r}
mob_data %>% 
  group_by(Device.Model) %>%
  summarise(battery_drain_time = (round(mean(Battery.Drain..mAh.day.), digits=0))) %>%
  arrange(desc(battery_drain_time))

ggplot(mob_data, aes(x=Device.Model, y=Battery.Drain..mAh.day.)) +
 geom_bar(stat = "identity", width=0.4)
```

